
---
title: "About Me"
author: "Edgar Pon"
date: "19/10/2021"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---


```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(gapminder)  # gapminder dataset
library(here)
#install.packages(janitor)
library(janitor)
```
`gapminder` country comparison

In this analysis, I have the `gapminder` dataset that has data on life expectancy, population, and GDP per capita for 142 countries from 1952 to 2007. To get a better look, I used glimpse function to see the variable names, variable types, etc., 
I also want to have a look at the first 20 rows of data.

```{r}
glimpse(gapminder)

head(gapminder, 20) # look at the first 20 rows of the dataframe

```

I have created the `country_data`for Portugal and `continent_data` for Europe with the code below.

```{r}
country_data <- gapminder %>% 
            filter(country == "Portugal") 

continent_data <- gapminder %>% 
            filter(continent == "Europe")
```

First, I plot life expectancy over the years for Portugal.

```{r, lifeExp_one_country}
plot1 <- ggplot(data = country_data, mapping = aes(x = year, y = lifeExp))+
geom_point() +
geom_smooth(se = FALSE)+
NULL 

plot1
```

And use `labs()` function to add an informative title to the plot.

```{r, lifeExp_one_country_with_label}
plot1<- plot1 +
labs(title = "Portugal Life Expectancy",
x = "Year",
y = "Life Expectancy") +
NULL


plot1
```

Next, I produce the plot for the european continent.

```{r lifeExp_one_continent}
ggplot(data=continent_data, mapping = aes(x = year , y = lifeExp , colour= country, group =country))+
geom_point() + 
geom_smooth(se = FALSE) +
NULL
```

Finally, I create graphs for all the continents and plot the life expectancy over time.

```{r lifeExp_facet_by_continent}
ggplot(data = gapminder , mapping = aes(x = year , y = lifeExp , colour= continent))+
geom_point() + 
geom_smooth(se = FALSE) +
facet_wrap(~continent) +
theme(legend.position="none") + #remove all legends
NULL
```

After a thorough analysis, I came to the following conclusion.

> Type your answer after this blockquote.

These graphs can be related to the evolution of demographic transition. One of the implications is that as economies develop and overall life quality improves, life expectancy tends to increase. All continents witnessed an increase in life expectancy, albeit at different growth rates and different age structures. Starting with the oceania and european continents, the most developed continents, the average life expectancy is the highest when compared to the other 3 continents. This result is expected due to improvements in healthcare and technology occurring in industrialized countries first. Moreover, in american and asian continents, a similar pattern is observable. However, as most of the countries in these two continents were not yet developed, the growth rate in life expectancy is expected to be higher than the rate in the already developed countries. Lastly, the african continent is the last continent to start the process of industrialization, but when compared to other continents, the growth rate of life expectancy seems to have stagnated in recent years, largely due to political failures, famine and lack of resources to improve living conditions. 
